|

### 2021-12-16 08-Discussion.Rmd
Proofreading edits

In __Chapter \@ref(sys-rev-tools-heading)__, I introduced the core concepts considered in this thesis and framed the research presented in relation to a theoretical framework of evidence synthesis.

@@

In __Chapter \@ref(sys-rev-methods-heading)__, I described the methods of a comprehensive systematic review of all available evidence on the association of both lipids and lipid-regulating agents with dementia outcomes. The results of this analysis are then presented in __Chapter \@ref(sys-rev-results-heading)__.

@@

As the results of the different components of this thesis are discussed in detail as part of the triangulation analysis (see Section \@ref(qual-tri)), here I briefly summarise the key findings.

Overall, I did not identify a consistent effect of any blood lipid fraction (total cholesterol, LDL cholesterol, HDL cholesterol or triglycerides) on any dementia outcome (all-cause dementia, Alzheimer's disease or vascular dementia). While published non-randomised studies provided evidence for a protective effect of lipid-lowering via statin use on Alzheimer's disease and the IPD analysis suggested a harmful effect of raised triglycerides on vascular dementia, these effects were not maintained when incorporated along with other sources of evidence in a quantitative triangulation framework. Similarly, the IPD analysis did not provide evidence for an interaction of patient characteristics (age/sex) with the association of blood lipids on dementia outcomes, other than a slight protective effect of male gender on the LDL-c/all-cause dementia relationship. However, this analysis was limited by the low number of cohorts analysed.

@@

Since the searches underpinning the systematic review described in this thesis were performed (July 2019), further studies on this topic have been performed. Two of these, notable for the analysis of large UK cohorts, are discussed here.

The first is a large-scale analysis of 953,635 patients in the CPRD, followed from their first LDL-c measurement.[@iwagami2021] The study found evidence for a slightly increased risk of dementia with higher LDL-c levels at mid-life, driven by the finding for the Alzheimer's disease subgroup. These results are consistent with the findings presented in this thesis, particularly as this study did not have access to _Apo_$\mathcal{E}4$ status and so could not adjust for confounding by this variable. _Apo_$\mathcal{E}4$ is a key risk factor for both Alzheimer's disease and raised LDL-c levels, <!-- TODO CITATION NEEDED --> and so could explain the harmful association observed. This is supported by a simulation analysis reported within this study, which used published information to predict that _Apo_$\mathcal{E}4$ would induce a RR of 1.09 per 1-SD increase in LDL-c in the absence of a true causal effect.

The second analysis examined 502,226 participants in the UK Biobank, a large population-based prospective cohort study.[@gong2021; @sudlow2015] The analysis found no association between lipids measured at mid-life and dementia outcomes, again a comparable finding to the results of this thesis. However, similar to the study discussed above, this analysis did not have access to genetic data and so may be subject to residual confounding by _Apo_$\mathcal{E}4$ genotype.

@@

It may well prove difficult to assess if any observed trends are due to a true effect or to confounding or ecological biases. However, because dementia outcomes will be monitored anyway for incidence/prevalence estimation and planning of services/care, this approach could provide a ready source of further evidence with a distinct bias structure to the approaches considered in this thesis.

@@

In terms of visualisation of bias, I have made a number of methodological contributions. I created a well-received R package to create paired forest plots (see Appendix \@ref(appendix-robvis)), and built on this to enable creation of the bias direction plots illustrated in Chapter \@ref(tri-heading).

Finally, I made several novel contributions around the incorporation of bias into analyses. These were intended to encourage authors of reviews to actively consider the impact of bias in their synthesis, something that empirical evidence suggests happens infrequently at present.[@katikireddi2015] The new methods introduced by this thesis range from the forest plots stratified by overall risk of bias level presented in Chapter \@ref(sys-rev-results-heading) to the more advanced triangulation framework presented in Chapter \@ref(tri-heading).

@@

The proposed quantitative triangulation framework represents the final novel contribution of this thesis to evidence synthesis methodology. Using pre-specified adjustment distributions mapped to the results of the domain-based risk-of-bias and indirectness assessments, this approach provides a systematic way to account for biases/indirectness across studies, reducing the potential for differential adjustment based on knowledge of the results. This framework explicitly seeks to include both direct and indirect evidence, and separates the assessment of bias/indirectness in each result from the definition of distributions representing the impact of bias/indirectness.

<!-- TODO Double check last paragraph! -->

@@

Finally, while the framework proposed here could (fairly) be criticised over the validity of the prior distributions of bias/indirectness chosen, the assumptions made about the impact of bias in this thesis are no stronger than those made when synthesising effect estimates as though they were unbiased. The proposed framework at the very least recognises the uncertainty introduced by bias/indirectness and decreases the precision of the overall effect estimate on this basis.

@@

There are several strengths to this thesis as a whole. Specifically, the identification and triangulation of evidence across study designs and publication status sets this thesis apart from previous syntheses of the evidence on this topic. Additionally, this thesis has also produced new evidence on the association of blood lipids with dementia outcomes in previously unanalysed cohorts, and provided additional information on a previously under-studied relationship (lipids/LRA and vascular dementia).

A further strength of this thesis is the production of software to support the novel evidence synthesis techniques used. Extensive documentation has been written to guide users through usage of the tools and example data are provided, enabling future researchers to readily apply the methods described. Finally, substantial effort has been invested to make the research documented in this thesis as reproducible as possible - this is documented more fully in Appendix \@ref(disc-repro-research).

@@

However, there are also several limitations to this thesis. In the first instance, as with many studies of dementia outcomes, a limiting factor in the interpretation of the results is the absence of a detailed pathological mechanism. It is possible that statins have a true effect via some pathway other than lipid-regulation, but without knowledge of the mechanism, the plausibility of this relationship cannot be assessed.

@@

A final weakness stems from the geographical focus of the data analysed in this thesis. All primary analysis presented drew on data from the UK (CPRD, CaPS, EPIC Norfolk, Whitehall II), while the majority of studies identified by the review were based in the Western world (Figure \@ref(fig:cohortLocations)). This, combined with the limited evidence for different ethnicities available even within high income settings, may limit the generalisability of the results presented to different populations.

@@

The proposal for the program of research underpinning this thesis was created in advance of starting my PhD, as a detailed plan was required in order to secure the funding that supported me through my studies. In hindsight, the decision to take a broad approach to the inclusion of dementia outcomes resulted in a larger workload than anticipated, particularly when also considering evidence from multiple different study designs. Additionally, as dementia subtypes likely have very different aetiological pathways, it may have been better to focus on a single subtype (e.g. Alzheimer's disease) when considering the causal effects of lipids. Similarly, attempting to undertake a full individual participant data meta-analysis as a single part of a larger program of research was overly ambitious, as data cleaning and harmonisation for just three cohorts turned out to be a substantial undertaking.

@@

### Generation of new RCT evidence

It is customary at this point in the discussion of results based (primarily) on observational data to recommend that a large-scale randomised controlled trial be performed to assess the effect of the proposed intervention. Given the costs and logistical challenges associated with trials examining outcomes with long prodromal periods such as dementia, a more efficient approach is offered by the opportunistic post-trial follow-up of existing RCTs of lipid-regulating agents.

This approach has already been used to assess long-term (11-20 years) safety outcomes of statin use, using participant recontact[@group2011] and linkage with electronic health records to identify events.[@ford2016] A similar approach could be used to assess the impact of randomisation to statins at midlife on subsequent risk of dementia outcomes. Use of data linkage would represent the most cost-efficient approach,[@llewellyn-bennett2018] though as noted in Chapter \@ref(cprd-analysis-heading), there is the potential for non-differential misclassification when defining dementia outcomes using EHRs.

@@

However, it seems unlikely that further analysis of observational data alone will be sufficient to investigate this outcome, as indicated by the presence of strong confounding by indication in the analysis of statins and vascular dementia (Chapter \@ref(cprd-analysis-heading)). Even using new methods, such as a target trial emulation approach, previous analyses have not replicated the known protective effect of statins on coronary heart disease identified by RCTs (see Section \@ref(cprd-confounding-by-ind)).[@danaei2013] Similarly, available Mendelian randomisation studies are limited to a single one-sample analysis of HMGCR variants as a proxy for statin treatment. This study suggested a protective effect of lipid-lowering on vascular dementia, though there was a moderate risk of bias in this analysis and the effect was attenuated to the null when including a wider range of lipid-lowering variants.

Taken together, these points reinforce the need for a large-scale GWAS of vascular dementia, similar to those currently available for Alzheimer's disease, to identify associated loci that future two-sample Mendelian randomisation studies could exploit. While GWAS of this outcome are likely to be methodologically challenging, given the difficulty in diagnosing "pure" vascular dementia, it would be a worthwhile endeavour and allow for the assessment of genetically driven risk factors beyond those considered in this thesis.

@@

The `medrxivr` tool enables programmatic searching of health-related preprints, and future work could incorporate the tool, along with software to search other literature sources, into an automated searching pipeline as part of a living systematic review.[@elliott2014] In addition, the ready access to preprint data afforded by the tool will enable future meta-epidemiological studies to examine factors which may influence eventual publication (for example,  significant result, geographical location/gender/career stage of first author, or some other factor). Finally, the tool will allow future research to assess the impact of peer review and editorial guidelines on manuscripts through comparison of the same paper at two stages (preprinted versus published). This approach was used to explore the impact of editorial polices on open data sharing, as discussed in Chapter \@ref(sys-rev-tools-heading), but could be applied to other aspects of the publication process.

>>>>

In __Chapter \@ref(background-heading)__, I introduced the core concepts considered in this thesis and framed the research presented in relation to a theoretical framework of evidence synthesis.

@@

In __Chapter \@ref(sys-rev-methods-heading)__, I described the methods of a comprehensive systematic review of all available evidence on the association of both lipid levels and lipid-regulating agents with dementia outcomes. The results of this analysis are then presented in __Chapter \@ref(sys-rev-results-heading)__.

@@

Given that the results of the different components of this thesis are discussed in detail as part of the triangulation analysis (see Section \@ref(qual-tri)), here I briefly summarise the key findings.

Overall, I did not identify a consistent effect of any blood lipid fraction (total cholesterol, LDL cholesterol, HDL cholesterol or triglycerides) on any dementia outcome (all-cause dementia, Alzheimer's disease or vascular dementia). While published non-randomised studies provided evidence for a protective effect of lipid-lowering via statin use on Alzheimer's disease and the IPD analysis suggested a harmful effect of raised triglycerides on vascular dementia, these effects were not maintained when incorporated along with other sources of evidence in a quantitative triangulation framework. Similarly, the IPD analysis did not provide evidence for an interaction of patient characteristics (age/sex) with the association of blood lipids and dementia outcomes, other than a slight protective effect of male gender on the LDL-c/all-cause dementia relationship. However, this analysis was limited by the low number of cohorts analysed.

@@

Since the searches underpinning the systematic review described in this thesis were performed (July 2019), further studies on this topic have been published. Two of these, notable for the analysis of large UK cohorts, are discussed here.

The first is a large-scale analysis of 953,635 patients in the CPRD, followed from their first LDL-c measurement.[@iwagami2021] The study found evidence for a slightly increased risk of dementia with higher LDL-c levels at mid-life, driven by the finding for the Alzheimer's disease subgroup. These results are consistent with the findings presented in this thesis, particularly as this study did not have access to _Apo_$\mathcal{E}4$ status and so could not adjust for confounding by this variable. _Apo_$\mathcal{E}4$ is a key risk factor for both Alzheimer's disease and raised LDL-c levels, <!-- TODO CITATION NEEDED --> and so could explain the harmful association observed. This is supported by a simulation analysis reported within this study, which used published information to predict that _Apo_$\mathcal{E}4$ would induce a risk ratio of 1.09 per 1-SD increase in LDL-c in the absence of a true causal effect.

The second analysis examined 502,226 participants in the UK Biobank, a large population-based prospective cohort study.[@gong2021; @sudlow2015] The analysis found no association between lipids measured at mid-life and dementia outcomes. This finding is again comparable to the results of this thesis. However, similar to the study discussed above, this analysis did not have access to genetic data and so may be subject to residual confounding by _Apo_$\mathcal{E}4$ genotype.

@@

It may well prove difficult to assess if any observed trends are due to a true effect or to confounding or ecological biases. However, because dementia outcomes will be monitored anyway for incidence/prevalence estimation and planning of services/care, this approach could provide a ready source of further evidence with a distinct bias structure to the approaches already considered in this thesis.

@@

In terms of visualisation of bias, I have made a number of methodological contributions. I built a well-received R package to create paired forest plots (see Appendix \@ref(appendix-robvis)), and developed this further to enable users to produce the bias direction plots illustrated in Chapter \@ref(tri-heading).

Finally, I made several novel contributions around the incorporation of bias into analyses. These were intended to encourage authors of reviews to actively consider the impact of bias in their syntheses, something that empirical evidence suggests happens infrequently at present.[@katikireddi2015] The new methods introduced by this thesis range from the forest plots stratified by overall risk of bias level presented in Chapter \@ref(sys-rev-results-heading) to the more advanced triangulation framework presented in Chapter \@ref(tri-heading).

@@

The proposed quantitative triangulation framework represents the final novel contribution of this thesis to evidence synthesis methodology. Using pre-specified adjustment distributions mapped to the results of the domain-based risk-of-bias and indirectness assessments, this approach provides a systematic way to account for biases/indirectness across studies. In addition, it explicitly seeks to include both direct and indirect evidence, and separates the assessment of bias/indirectness in each result from the definition of distributions representing the impact of bias/indirectness.

@@

Finally, the framework proposed here could (fairly) be criticised over the validity of the prior distributions of bias/indirectness chosen. However, the assumptions made about the impact of bias in this thesis are no stronger than those made when synthesising effect estimates as though they were unbiased. The proposed framework at the very least recognises the uncertainty introduced by bias/indirectness and decreases the precision of the overall effect estimate on this basis.

@@

There are four key strengths to this thesis as a whole. Firstly, the identification and triangulation of evidence across study designs and publication status sets this thesis apart from previous syntheses of the evidence on this topic.

Additionally, this thesis has also produced new evidence on the association of blood lipids with dementia outcomes in previously unanalysed cohorts, and provided additional information on a previously under-studied relationship (lipids/LRA and vascular dementia).

A further strength of this thesis is the production of software to support the novel evidence synthesis techniques used. Extensive documentation has been written to guide users through usage of the tools and example data are provided, enabling future researchers to readily apply the methods described.

Finally, substantial effort has been invested to make the research documented in this thesis as reproducible as possible - this is documented more fully in Appendix \@ref(disc-repro-research).

@@

However, there are also four key limitations to this thesis. As with many studies of dementia outcomes, a limiting factor in the interpretation of the results is the absence of a detailed pathological mechanism. It is possible that statins have a true effect on dementia risk via some pathway other than lipid-regulation, but without knowledge of the mechanism, the plausibility of this relationship cannot be assessed.

@@

A final weakness stems from the geographical focus of the data analysed in this thesis. All primary analysis presented drew on data from the UK (CPRD, CaPS, EPIC Norfolk, Whitehall II), while the majority of studies identified by the review were based in westernised countries (Figure \@ref(fig:cohortLocations)). This, combined with the limited evidence for different ethnicities available even within high income settings, may limit the generalisability of the results presented to different populations.

@@

The proposal for the program of research underpinning this thesis was created in advance of starting my PhD because a detailed plan was required in order to secure the funding that supported me through my studies. In hindsight, the decision to take a broad approach to the inclusion of dementia outcomes resulted in a larger workload than anticipated, particularly when also considering evidence from multiple different study designs. Additionally, as dementia subtypes likely have very different aetiological pathways, it may have been better to focus on a single subtype (e.g. Alzheimer's disease) when considering the causal effects of lipids. Similarly, attempting to undertake a full individual participant data meta-analysis as a single part of a larger program of research was overly ambitious, given that data cleaning and harmonisation for just three cohorts turned out to be a substantial undertaking.

@@

### Generation of new evidence from RCTs

It is customary at this point in the discussion of results based (primarily) on observational data to recommend that a large-scale randomised controlled trial (RCT) be performed to assess the effect of the proposed intervention. Given the costs and logistical challenges associated with trials examining outcomes with long prodromal periods such as dementia, a more efficient approach is offered by the opportunistic post-trial follow-up of existing RCTs of lipid-regulating agents.

This approach has already been used to assess long-term (11-20 years) safety outcomes of statin use, using participant recontact[@group2011] and linkage with electronic health records to identify events.[@ford2016] A similar approach could be used to assess the impact of randomisation to statins at midlife on subsequent risk of dementia outcomes. Use of data linkage would represent the most cost-efficient approach,[@llewellyn-bennett2018] though as noted in Chapter \@ref(cprd-analysis-heading), there is the potential for non-differential misclassification when defining dementia outcomes using electonic health records.

@@

However, it seems unlikely that further analysis of observational data alone will be sufficient to investigate this outcome, given the evidence of strong confounding by indication presented by the analysis of statins and vascular dementia (Chapter \@ref(cprd-analysis-heading)). Even using cutting-edge methods, such as a target trial emulation approach, previous analyses have not replicated the known protective effect of statins on coronary heart disease identified by RCTs (see Section \@ref(cprd-confounding-by-ind)).[@danaei2013] Similarly, available Mendelian randomisation studies are limited to a single one-sample analysis of HMGCR variants as a proxy for statin treatment. This study suggested a protective effect of lipid-lowering on vascular dementia, although there was a moderate risk of bias in this analysis and the effect was attenuated to the null when including a wider range of lipid-lowering variants.

Taken together, these points reinforce the need for a large-scale genome-wide association study (GWAS) of vascular dementia, similar to those currently available for Alzheimer's disease, to identify genetic variants associated with this outcome that future two-sample Mendelian randomisation studies could exploit. While GWAS of this outcome are likely to be methodologically challenging, given the difficulty in diagnosing "pure" vascular dementia, it would be a worthwhile endeavour and allow for the assessment of genetically driven risk factors beyond those considered in this thesis.

@@

The `medrxivr` tool enables programmatic searching of health-related preprints, and future work could incorporate the tool, along with software to search other literature sources, into an automated searching pipeline as part of a living systematic review.[@elliott2014] In addition, the ready access to preprint data afforded by the tool will enable future meta-epidemiological studies to examine factors which may influence eventual publication (for example,  significant results, geographical location/gender/career stage of first author, or some other factor). Finally, the tool will allow future research to assess the impact of peer review and editorial guidelines on manuscripts through comparison of the same paper at two stages (preprinted versus published). This approach was used to explore the impact of editorial polices on open data sharing, as discussed in Chapter \@ref(sys-rev-tools-heading), but could be applied to other aspects of the publication process.
